---
title: "Class09"
author: "Megan Truong"
date: "2/4/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K- means clustering

the main l-means function inR is called kmeans()
rev: reverses order of function

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)

```

Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km

# Cluster size
km$size
# Cluster
km$cluster
#cluster center
length(km$cluster)
table(km$cluster)

```
 
```{r}
plot(x, col=km$cluster)
points(km$centers, col = "blue", pch = 15, cex =3)
```
 
 
 
 ## Hierarchial clustering in R
 
 The main Hierarchical clustering function in R is called hclust() 
    Important: have to calculate the distance matrix deom of input data     before calling hclus()
  
```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc 
```
```{r}
plot(hc)
abline(h=6, col ="red", lty =2)
abline(h=4, col="blue")
```

To get cluster membership- vector need to cut the tree at a certian height to yield separate cluster branches
```{r}
cutree(hc, h=6)
```

```{r}
gp4 <- cutree(hc, h=4)
table(gp4)
```


## More

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
 Boundaries of clusters are not always clear...need to cut
```{r}
hc <- hclust(dist(x))
hc
```

```{r}
plot(hc)
abline(h=1.7, col = "red")
```
Cut Tree
```{r}
grps <- cutree(hc, k =3)
grps
table(grps)
plot(x, col = grps)
```

## PCS of Uk food data

```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x
```

```{r}
# number of rows and columns
nrow(x)
ncol(x)
```

Making plots to explore the data

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```


```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```


A "pairs" plot can be helpful when we have small datasets like this one but  often we  are dealing with  data that is too large for this approach

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
attributes(pca)
```


```{r}
plot(pca$x[,1], pca$x[,2])
```


```{r}
summary(pca)
```





```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x),
     col=c("black", "red", "blue", "green"))
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

