---
title: "Transcriptomics and the analysis of RNA-Seq data"
author: "Megan Truong"
date: "2/20/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data for this hands-on session comes from a published RNA-seq experiment where airway smooth muscle cells were treated with dexamethasone, a synthetic glucocorticoid steroid with anti-inflammatory effects (Himes et al. 2014).

Glucocorticoids are used, for example, by people with asthma to reduce inflammation of the airways. The anti-inflammatory effects on airway smooth muscle (ASM) cells has been known for some time but the underlying molecular mechanisms are unclear.


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```


How many genes are we delaing with?

```{r}
nrow(counts)
```

To calculate the average counts per gene for control sampless and then for drug treated ones. Then compare to see if there is a difference between control and treated samples for a given gene.


This is metadata info that tells where our control and treated samples are 
```{r}
metadata
```


The "dex" column tells us which is control or treated
```{r}
metadata$dex
```


To see which are control, will tell you true or false
```{r}
metadata$dex == "control"
```


To see which of them is a control by id use []. If you want to access control only rows in metadata
```{r}
control <- metadata[metadata$dex == "control", ]
#metadata$id[metadata$dex == "control"]

counts[ , control$id]

```


Now use $id of control data to access control columns in counts and calculate the average
```{r}
control.mean <- rowSums(counts[ , control$id]) / 4
# divide by four bc there are 4 control samples or divide by nrow(control)

names(control.mean) <- counts$ensgene
# put the names of the genes back on


control.mean
```


Now do the treated ones
```{r}
treated <- metadata[metadata$dex == "treated", ]

counts[ , treated$id]
```

```{r}
treated.mean <- rowSums(counts[ , treated$id]) / nrow(treated)

treated.mean
```



For bookkeeping purposes let's combine the meancount data:

```{r}
meancounts <- data.frame(control.mean, treated.mean)
head(meancounts)
```



Let's plot

```{r}
plot(meancounts)
```


Let's make this plot log-log plot
```{r}
plot(log2(meancounts))
```

We can find candidate differentially expressed genes by looking for genes with a large change between control and dex-treated samples. We usually look at the log2 of the fold change, because this has better mathematical properties.

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the head() or the View() function for example.

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)

```



Remove any gene that has zero average counts

```{r}
x <- data.frame(c(1,1,10,0), c(0,1,10,0))
x
x == 0
which(x == 0, arr.ind = TRUE)[, "row"]
```



```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

How many genes left?
```{r}
nrow(mycounts)
```


A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```


How many upregulated genes at this threshold
```{r}
sum(up.ind)
```

How many downregulated genes at this threshold?
```{r}
sum(down.ind)
```


## DESeq2

```{r}
library(DESeq2)
```

Setup input for DESeq by specifiying countdata, metadata, and the experiment design (i.e which comparison we are interested in)
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
```

```{r}
plot(res$log2FoldChange, res$padj)
```


```{r}
plot(res$log2FoldChange, log(res$padj))
```



```{r}
plot(res$log2FoldChange, -log(res$padj))
```

```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

#Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```

