---
title: "Class 12: Bioinformatics pt. 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Here we inspect the types of structures in the main database for 3D biomolecular data - the PDB. 


NOTE: The “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type” 

> Q1. Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
# read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)

# calculate percentage values for each 'method'
# total number of structures in data set
sum(stats$Total)

ans <- (stats$Total / sum(stats$Total)) * 100
names(ans) <- rownames(stats)
round(ans, 2)

```


> Q2. Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
protein <- round(sum(stats$Proteins) / sum(stats$Total) * 100, 2)
protein

```



> Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?





