---
title: 'Class 06: R Functions'
author: "Megan Truong"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

## Level 2 Heading

### Level 3 Heading

plot(1:5, col="blue) will run as text bc didn't insert r code

```{r}
#this is a silly plot
plot(1:5)
```

read.table("Your File Name", header = TRUE/FALSE, sep="\t"[tab] or comma or space)


More about file import (i.e. reading files into R). The main read function in base R is 'read.table()'. 

```{r}
T1 <- read.table("bimm 143 class 6 test 1.txt")
```

```{r}
T1 <- read.table("bimm 143 class 6 test 1.txt", sep = ",", header = TRUE)
```

or use 'read.cvs()' which has the arguments we want in this case already
```{r}
T1 <- read.csv("bimm 143 class 6 test 1.txt")
```


Example 2:
```{r}
T2 <- read.table("bimm 143 class 6 test 2.txt", header = TRUE, sep = "$")
```

Example 3: 
```{r}
read.table("bimm 143 class 6 test 3.txt")
```


```{r}
add <- function(x, y) {
  # Sum the input x and y
  x + y
}
```

```{r}
add(7, 3)
```


```{r}
add(c(1,2,3), 4)
```

What does this do?
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))

df$a <- rescale(df$sa)

Simplify:
x <- (x - **min(x)**) / (max(x)) - **min(x)**)
xmin <- min(x)
x <- (x - xmin) / (max(x)) - xmin)

What is the 'range()' function?
```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)
```


```{r}
range(x)
```

This is our second function
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale(x)
```

more testing
```{r}
rescale(1:10)
```

```{r}
rescale( c(1,2,NA,3,10) )
```

```{r}
x <- c(1,2,NA,3,10)
rng <-range(x)
rng
```

How do we argue with it?
```{r}
rng <-range(x, na.rm = TRUE)
rng
```


```{r}

rescale2 <- function(x) {
rng <- range(x, na.rm = TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}

```


```{r}
rescale2( c(1,2,NA,3,10) )
```


```{r}
rescale <- function(x, na.rm=TRUE, plot=FALSE) {
  
rng <-range(x, na.rm=na.rm)
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 print("is it me you are looking for?")
 
if(plot) {
   print("Don't sing agian please!")
plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```


```{r}
rescale <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 return(answer)
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```




Install only 1 time, library is when you need it
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```

*Q1. What type of object is returned from the read.pdb() function?*
A data table from a protein data bank file (brom bio3d)
```{r}
class(s1)
```
```{r}
str(s1)
```

```{r}
s1$seqres
```

```{r}
aa321(s1$seqres)
```

```{r}
s1$atom
```


*Q2. What does the trim.pdb() function do?*
Trims the bigger PBD that contains subset of atoms; makes it a smaller set


*Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?*



*Q4. What would be a better plot to compare across the different proteins?*
```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ= "l", col= "blue", lwd = 2)
points(s3.b, typ="l", col = "red", lwd = 2)
```



*Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a resulting dendrogram plot.*

```{r}
hc <- hclust(dist(rbind(s1.b, s2.b, s3.b)))
```

```{r}
plot(hc)
```


*Q6. How would you generalize the original code above to work with any set of input protein structures?*
The protein structure
x <- the protein structure












