---
title: 'Class 6: R Fucntion'
author: "Megan Truong"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## My level 2 heading

This is an *R Markdown* document. **COOL**

This is just old text that renders as text

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:


Use Insert button and choose R to write R code
```{r}
plot(1:5, typ="l", col="blue")
```
click lil green arrow to show chunk

or use command + option + i
```{r}

```
 


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
