---
title: "Class13"
author: "Megan Truong"
date: "2/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hands on Section

## Dowlaod and clean HIV-Pr receptor structure 

Here we downloaded the 1HSG PDB structure from the online database, remove the bound drug from the Merk and write out a porotien only and ligand only new set of PDB files 

You can copy and paste and use it on any protien just change the name!!

Just use bio2d **package functions**: get.pdb(), read.pdb(), atom.select() and write.pdb().

```{r}
library(bio3d)

# Download PDB file: where 1hsg is any sturcture/protein you want to search
file.name <- get.pdb("1hsg.pdb")

pdb <- read.pdb(file.name)

pdb
```

Select protein only
```{r}
prot <- atom.select(pdb, "protein", value = TRUE)

write.pdb(prot, file = "prot") 

prot
```


Select and write out ligand only
```{r}
lig <- atom.select(pdb, "ligand", value = TRUE)

write.pdb(lig, file = "lig")

lig

```


>Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 














