---
title: "Class_07"
author: "Megan Truong"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

fname <- function(arg1, arg2) {
}


Theese are functions here
```{r}
function(x, na.rm=TRUE, plot=FALSE, ...) {
  # Our rescale function from the end of lecture 9

 if( !is.numeric(x) ) {
 stop("Input x should be numeric", call.=FALSE)
 }
   
  if(na.rm) {
    rng <-range(x, na.rm=TRUE)
  } else {
    rng <-range(x)
  }

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  if(plot) { 
    plot(answer, ...) 
  }

  return(answer)
}

```


"warning()" break functions that will give warning and text; tolerable error

"stop()" stops running after warning for intolerable error


```{r}
# Lets define an example x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.numeric(x)
```

```{r}
is.na(x)
```

Compare x and y
```{r}
is.na(x) & is.na(y)
```

```{r}
is_both_na <- function(x,y) {
  if(length(x) != length(y))
    stop("your inputs aren't the same length")

  return(na.number)
}




```


```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
z <- c(1,1)


```

```{r}
rescale_new  <- function(x,y) {
  if(length(x) != length(y))
    stop("your inputs aren't the same length", call = FALSE)

  na.in.both <- ( is.na(x) & is.na(y) )
  na.number <- sum(na.in.both)
  na.which <- which(na.in.both)
  
  message("Found", na.number, "nA's at position(s):",
          paste(na.which, collapse = ","))
  
  return(list(number = na.number, which = na.which))
}


rescale_new(x, y)

```

```{r}
# Start with a simple version of the problem
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
  exp=c(2,1,1),
  stringsAsFactors=FALSE)
df2 <- data.frame(IDs=c("gene2", "gene4", "gene3", "gene5"),
  exp=c(-2, NA, 1, 2),
  stringsAsFactors=FALSE)

# Simplify further to single vectors
x <- df1$IDs
y <- df2$IDs

x
y

intersect(x,y)

x %in% y
x[x %in% y]
y[y %in% x]

cbind(x[x %in% y], y[y %in% x])
```


Make a function: highlight what you want to make a function, click Code in bar and "Extract Function"

```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y], y[y %in% x])
}


x <- df1$IDs
y <- df2$IDs
gene_intersect(x,y)

gene_intersect2 <- function(df1, df2) {
   cbind(df1[df1$IDs %in% df2$IDs, ], 
         df2[df2$IDs %in% df1$IDs, ], "exp")
  
}

gene_intersect2(df1, df2)


```




```{r}
gene_intersect3 <- function(df1, df2, gene.colname="IDs") {

 cbind( df1[ df1[,gene.colname] %in%
 df2[,gene.colname], ],
 exp2=df2[ df2[,gene.colname] %in%
 df1[,gene.colname], "exp"] )

}
 
# These are all equivalent functions
df1$IDs
#df1[,"IDs"]
gene.colname = "IDs"
#df1[ , gene.colname]
 
colnames(df1) = c("IDs", "exp")
df1
colnames(df2) = c("IDs", "exp")


x <- df1$IDs
y <- df2$IDs
gene_intersect(x,y)


gene_intersect3(df1, df2, gene.colname = "IDs")
  
```




```{r}
# Improve by simplifying for for human consumption
gene_intersect4 <- function(df1, df2, gene.colname="IDs") {

 df1.name <- df1[,gene.colname]
 df2.name <- df2[,gene.colname]

 df1.inds <- df1.name %in% df2.name
 df2.inds <- df2.name %in% df1.name

 cbind( df1[ df1.inds, ],
 exp2=df2[ df2.inds, "exp"] )
}

gene_intersect4(df1, df2)
```




```{r}
library("ggplot2")

dim(iris)

class(iris)

head(iris)
```


```{r}
myplot <- ggplot(data = iris, aes(x = Sepal.length, y = Sepal.Width))
class(myplot)

ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point()
```



Change point size, color and shapes

```{r}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) + geom_point(aes(shape = Species), size = 3)
```


With fitted lines
```{r}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) + geom_point(aes(shape = Species), size = 3) + geom_smooth(method = "lm")
```


For a bar graph, alpha changes transparency
```{r}
ggplot(iris, aes(Sepal.Length, fill = Species)) + geom_bar(stat = "count", alpha = 0.7)
```

you can change way more but we ran out of time