---
title: "Class10"
author: "Megan Truong"
date: "2/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scaling
prcomp(x, scale = TRUE)


# Analysis of Human Breast Cancer Cells

First import data (read)

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

There are some things in this data set that we will ignore, including the ID and Diagnosis columns and weird last X column. We only want columns 3-32

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```

How many patients do we have data for?
```{r}
nrow(wisc.data)
```


How many Ms and Bs are in the diagnosis column?
```{r}
table(wisc.df$diagnosis)
```

How many variables/features in data are suffixed with _mean?

```{r}
x <- colnames(wisc.data)
# gives a list of which number columns have them. can  also do grep("_mean", colnames(wisc.data))
grep("_mean", x)

# gives list of names with pattern
grep("_mean", x, value = TRUE)

# for how long it is/ how many have the pattern
length((grep("_mean", x, value = TRUE)))
```



## Principal Component Analysis

Before we do PCA we need to think/consider whether we should SCALE our input

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.

The input variables have significantly different variances.

Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before.

```{r}
# column means
colMeans(wisc.data)

# column standard deviations
apply(wisc.data, 2, sd)

# Round numbers to  2 decimal places
round (apply(wisc.data, 2, sd), 2)
```


We need to set scale = TRUE

```{r}
# Perform pCA on wisc.data 
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

From your results, what proportion of the original variance is captured by the first principal components (PC1)?

The first PC captures 44.27% of the proportion of the original variance. 72.67% in the first 3 PCs.

Lets make some PC figures

```{r}
# this one ugly af "a hot mess"
biplot(wisc.pr)
```

We need our own PC1 vs PC2 plot colored by the expert diagnosis.

```{r}
attributes(wisc.pr)
```

```{r}
# basic ass plot
plot(wisc.pr$x[,1], wisc.pr$x[,2])
```


```{r}
# colored by diagnosis (malignant is red and benign is black)
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = wisc.df$diagnosis,  xlab = "PC1", ylab = "PC2")
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(h=0, col="gray", lty=2)
abline(v=0, col="gray", lty=2)
```



## Cluster in PC space

First let's see if we can cluster original data

kmeans(wisc.dat)
```{r}
wisc.hc <- hclust(dist(wisc.data))
plot(wisc.hc)

# lol that shit uglyyyy
```


Let's try to combine the results of PCA with clustering to make it look better. Let’s see if PCA improves or degrades the performance of hierarchical clustering.

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.

```{r}
wisc.pr.hclust <- hclust( dist(wisc.pr$x[,1:3]), method="ward.D2" )
plot(wisc.pr.hclust)
```




Let's cut it

```{r}
grps3 <- cutree(wisc.pr.hclust, k =3)
table(grps3)
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3)
```
Looks like:
  black and red correspond to malignant
  green corresponds to benign


We can use table() to compare $diagnosis vector with cluster results
```{r}
table(grps3, wisc.df$diagnosis)
```


## Prediction

We will use predict() to take our PCA model and new cancer cell data and project that data onto our PCA space

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

Use the predict function with previous PCA model and new data and draw the PCA plot again and add new data

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = wisc.df$diagnosis)
points(npc[,1], npc[,2], col ="blue", pch = 15, cex = 3)
text(npc[,1], npc[,2], labels = c(1,2), col = "white")
```



